Finite Time Bounds for Temporal Difference Learning with Function Approximation: Bugs in state-of-the-art results

### 2017-07-29 rebuttal.txt

3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? Why is our bound better? Also: line 275

5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise. Since the LSTD equation is not derived from a loss, we are puzzled about the comment about SGD. Mirror-prox (=GTD) comes closest to what we think SGD may look like in this setting.

Difference to state of the art: See above.
Upper bound no guidelines: Lower bound shows there will be no guidelines. The adaptive algorithm is meant to address this, but further work is needed on this for sure. We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.

Rev2: We can push the adaptive stepsize section to the appendix and provide proof sketch in the main body.

Rev 1: Markov noise is absent in the data-resampling setting (quite common in RL [4,10,15,16]), where the noise is i.i.d. Novelties are analysis of general LSA schemes with multiplicative noise for constant stepsize+ PR avg (is absent in prior literature; only SGD is known); a negative result: problem dependent stepsize in RL (with matching lower bound). In concluding that there is no novelty in the paper, we note that the review here fails to justify his position, which is unwarranted.

### 2017-07-29 rebuttal.txt

3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? Why is our bound better? Also: line 275

5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise. Since the LSTD equation is not derived from a loss, we are puzzled about the comment about SGD. Mirror-prox (=GTD) comes closest to what we think SGD may look like in this setting.

Difference to state of the art: See above.
Upper bound no guidelines: Lower bound shows there will be no guidelines. The adaptive algorithm is meant to address this, but further work is needed on this for sure. We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.

Reviewer 2: We can push the adaptive stepsize section to the appendix and provide proof sketch in the main body.

Reviewer 1: 
1) Markov noise is absent in the data-resampling setting (quite common in RL [4,10,15,16]), where the noise is i.i.d. 
2) Novelties are significant- i) Analysis of general LSA schemes with multiplicative noise for constant stepsize+ PR avg is new in literature. Prior results were known only for general LSA with additive noise and for SGD (a special case of LSA) ii) Matrices in general LSA schemes are not symmetric and prior results do not apply directly; the similarity transformation that we employ in the paper is needed to handle the non-symmetric case; this is a major step iii) We show a negative result: for general LSAs the stepsize cannot be chosen in a problem independent manner; we provide a lower bound to show that the negative result is not due to a weakness in our analysis iv) Better rates for GTD follows immediately due to our results. Our paper obviates the projection step in GTD; this is a major gain.
We would like to mention that we have discussed the significance of our contributions in the paper in the introduction as well as section 6. In concluding that there is no novelty in the paper, we are afraid that the reviewer here fails to justify his position which is unwarranted.

### 2017-07-29 rebuttal.txt

We thank the reviewers for their comments.
Reviewer 1:
Naturally, we are disappointed by the negative recommendation, which we think is perhaps the result of overlooking some facts:
1) Two justifications for the iid setting: The noise is iid in the data-resampling setting, which is quite common in RL [4,10,15,16]. Understanding the differences between the standard supervised setting and the RL setting should start with a setting with the least number of differences. We see that already the iid noise setting presents surprises (see below).
2) Novelties are significant:
i) Analysis of general LSA schemes with multiplicative noise for constant stepsize combined with PR averaging is new in literature. Prior results were known only for general LSA with additive noise and for SGD (a special case of LSA). The new analysis required significant departure from the previous analysis and produces different results.
ii) Matrices in general LSA schemes lack symmetry, hence prior results do not apply directly; the similarity transformation that we employ in the paper is needed to handle the non-symmetric case; this is a major step
iii) We show a negative result: for general LSAs the stepsize cannot be chosen in a problem independent manner; we provide a lower bound to show that the negative result is not due to a weakness in our analysis
iv) Our results give better than previous rates for GTD. This way, our paper obviates a highly problematic projection step in GTD introduced by[10]; this is a major gain. The projection step is problematic because the range of dual variables cannot be known a priori in GTD.
We would like to mention that we have explicitly discussed the significance of our contributions in the paper in the introduction as well as section 6. Hence, we think that the reviewer's position is unwarranted and lacks justification.

Reviewer 2:
We will be happy to push the adaptive stepsize section to the appendix and provide a proof sketch in the main body.

Reviewer 3:
1) Missing citation to Prashanth, Korda and Munos' ECML'14 paper.
In fact, Prashanth and Korda later published arxiv.org/pdf/1411.3224.pdf (ICML'15), which we thought was the "state-of-the-art" and we cited this paper. Strangely, this paper does not cite the ECML'14 paper and so we failed to find the ECML'14 paper and hence failed to cite it. Now, as to the actual contributions, unfortunately, the ECML'14 paper seems to have many problems. We believe that the stated results in it are incorrect and fixing the bugs would lead to an analysis and results identical to ours. The factual errors are severe: for the expected error bound, on page 16 in Section A.2 of the arxiv version, after Eq. (26) it is claimed that \bar A_n - (1-beta) mu I is positive definite. This is easily seen to be false, e.g, when n = 1, in which case \bar A_n is rank 1, asymmetric and has nothing to do with mu, the minimum eigenvalue of the empirical feature covariance matrix. This is not a minor problem: Fixing the proof requires arguments similar to ours. The high-probability bound is similarly incorrect: On page 14, the display in the bottom of the page is just incorrect. Again, fixing the proof requires extra conditions, leading to an analysis similar to ours. As a sidenote, we remark that even the ICML'15 seems to have problems: (Earlier) we have contacted the first author, Prashanth, who confirmed these problems.

2) Work of Bach and Moulines: We think this is a precursor to and is superseded by the papers by Défossez and Bach [2] (see page 5 in [2]) and that of Dieuleveut, Flammarion and Bach [3] (see page 2 in [3]), which we cite and compare to (page 6, lines 226-241). In fact, we think that this comparison is truly enlightening as it brings out in what way the LMS update is a uniquely special case of LSA.

2) Knowledge of minimum eigenvalue: Indeed, the stepsize dependence on a quantity that is most likely unknown in an application is highly problematic! Our intention if anything was to make this very clear. In fact, here is where our lower bound becomes crucial: We are showing that the dependence is not an artifact of the analysis, but a weakness of the approach. Shining light on this dependence makes the paper uniquely valuable. We realize that there is a tendency in the community to prefer constructive and positive results, but we think that negative results like ours are also extremely useful: there is much to be learned from knowing what does not work (and why). Gaining this level of understanding is only possible by a careful theoretical analysis like ours. The importance of our result is further elevated by that the topic of RL/TD/.. is rapidly gaining popularity, and that numerous works attempt to transport results from supervised learning to this setting, overlooking the extra difficulties (e.g., see the papers by Prashanth et al.). Our paper may be valuable for both people solely interested in RL, but also for the fast growing community of researchers coming to RL with a supervised learning background.

3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? See above. Why is our bound better? Also: line 275

5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise (as recognized by many earlier, e.g., [4,10,15,16]). Since the LSTD equation is not derived from a loss, we are puzzled about the comment about SGD. Mirror-prox (=GTD) comes closest to what we think SGD may look like in this setting.

Difference to state of the art: See above.
Upper bound gives no guideline: Lower bound shows there will be no guidelines; an adaptive algorithm is necessary (contrasting LMS). Our adaptive algorithm is meant to address this, but further work is needed on this for sure. We think even without this further work, the work should be interesting for the community as outlined above (see also the response to Reviewer 1).

### 2017-07-29 issues.tex
updated issues
rebuttal? issues..

\documentclass{article}
\input{pack}
% if you need to pass options to natbib, use, e.g.:
 \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

@@

\title{Issues with Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control}

@@

\author{
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

\maketitle
In what follows we describe the issues with \citet{flstd}. The complete proofs of \cite{flstd} are available in the arxiv version in \cite{flstda}.

\section{Problems with the Proof}

\subsection{Expected Error Bound}

The proof of Theorem 1 can be found in section A.2, starting on page 16 of \cite{flstda}.
The proof up to Eq. (26) is correct (in the definition of $M_{n+1}$, $F_n$ should be $F$).
However, after this display, we are told that by (A4), which ensures that $\frac1T \sum_{t=1}^T \phi_t \phi_t^\top \succ \mu I$ for some $\mu>0$, $\bar A_n = \frac1n \sum_{i=1}^n \phi_i (\phi_i - \beta \phi_i')^\top$ is such that $\bar A_n - (1-\beta) \mu I $ is positive definite. 
Here, $n>0$ is an arbitrary index and we use the abbreviation $\phi_t = \phi(s_{i_t})$ and $\phi_t' = \phi(s_{i_t}')$.
\begin{quote}
In general, (A4) does \emph{not} imply that  $\bar A_n - (1-\beta) \mu I $ is positive definite!
\end{quote}
Take for example $n=1$. We would need that $\phi_1 (\phi_1 - \beta \phi_1')^\top - (1-\beta) \mu I$ is positive definite.
It is easy to construct examples where this is not true. (Note that the matrices involved are \emph{not} symmetric. Unfortunately, none of the two papers defines what is meant by positive definite in this case. We assume that the definition used is that a square, real-valued matrix $A$ is called positive definite if $x^\top A x\ge 0$ for any $x$ real-valued vector of appropriate dimension.)

In fact, we don't see whether this holds even when $n=T$ and we in fact suspect that this claim is false in full generality.

\section{Problems with the Proof: High Probability Bound}

The proof of the high probability bound starts on page 14, in section A.1.
The first problem happens in the display on the bottom of this page in the proof of Lemma 6.
Here, we are told that for $A = \phi \phi^\top - \beta \phi (\phi')^\top$ (we are dropping indices to remove clutter),
\begin{align*}
A^\top A = \norm{\phi}^2 \phi \phi^\top - \beta ( 2 - \norm{\phi}^2\beta ) \phi' (\phi')^\top\,,
\end{align*}
where $\norm{x}$ denotes the $2$-norm of $x$.
However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation gives:
\begin{align*}
A^\top A 
& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top \\
& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \\
& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\,,
\end{align*}
which does not match the previous display.
The terms that do not match are the linear-in-$\beta$ terms.
In the first display we have $ -2 \beta \phi' (\phi')^\top$, while in the bottom we have $-\beta (\phi' \phi^\top + \phi (\phi')^\top)$.

The first equality of their display states (in equivalent form) that 
\begin{align*}
A^\top A = \norm{\phi}^2 ( \phi \phi^\top - 2\beta \phi (\phi')^\top +\beta^2 \phi' (\phi')^\top )\,.
\end{align*}
We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.

We also have a hard time following the rest of the proof of this lemma.
In fact, we have a simple proof for a result similar to stated in this lemma.
This works by observing that one needs to 
bound the Lipschitz factor of $\theta \mapsto \bar T_n \dots \bar T_{t+1} ( (I-\gamma_t A_t)\theta + b_t )$, where
$A_t = \phi_t (\phi_t -\beta \phi_t')^\top$, $b_t = r_{i_t} \phi_t$ and
 $\bar T_t(\theta) = (I-\gamma_t \bar A_T) \theta + \bar b_T$ (we assume that the stepsize sequence is deterministic).
 Clearly, this Lipschitz factor can be bounded by $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 \norm{I-\gamma_t A_t}_2$
 (the composition of Lipschitz maps is Lipschitz with a factor that is the product of the individual maps' Lipschitz factors). 
Invoking (A3) then indeed gives some version of Lemma 7.
By making an appropriate assumption (similar to those used in our paper), 
$\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.

\section{Bounded Linear Space Assumption}
\citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}). This assumption is critical in establishing Lemma~7 (see pages 15 and 16, \cite{flstda}).
We suspect the following shortcomings with assumption \textbf{$(A3)$}
\begin{itemize}
\item The assumption is stated in a somewhat sloppy fashion. We take the authors meant to say that $\sup_n \norm{\Phi \theta_n}_\infty<+\infty$ holds almost surely. This seems like a strong assumption! Would not one of the main question be whether this assumption even holds? We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
\item One possibility would be to modify the algorithm by adding a projection step to guarantee boundedness.
It is still unclear to us whether this alone would ensure convergence of the error to zero. In any case, the analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).
%For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly). Then would the recursion unfold the same way as it does in Equation (26) of \cite{flstda} (or Equation (13) of \cite{flstd})? How are we to ensure decay/contraction (as shown in Equation (27) of \cite{flstda}) at each step (since projection might be non-contracting)?
%\item The recursion eventually (see Equation (26) of \cite{flstda}) has only \emph{additive noise} and not \emph{multiplicative noise} as analyzed in our submission. The additive noise is captured in the form of the $H_\beta^2$ term in Equation (28) of \cite{flstd}. This reduction to additive noise is possible due to \textbf{$(A3)$}.
\end{itemize}
\section{Relation between Covariance Matrix and $\bar{A}_T$ matrix}
\citet{flstda} assume positive definiteness of \textbf{$(A4)$} covariance matrix $\frac{1}{T}\Phi^\top_T \Phi_T$. However, unlike regression problems, in reinforcement learning problems what appears in the recursion (see Equation (6)) is not the covariance matrix, but a different matrix $\bar{A}_T=\frac{1}{T}\sum_{i=1}^T \phi(s_i)(\phi(s_i)-\beta\phi(s_i'))^\top$ defined in pages 2, 4 (below Equation (5)), 8 and 16 of \cite{flstda}. Usually, without a sampling assumption known as the `on-policy' case (see \cite{gtd} for a discussion on `on-policy' vs `off-policy') the eigenvalues of $\bar{A}_T$ cannot be guaranteed to have all positive real parts. While \citet{flstda} mention the `on-policy' sampling in the introduction, there is no explicit sampling assumption in the list of assumption. In fact, we doubt that the proposed algorithm will converge without extra assumption (as discussed above).
\if0
This is worrisome for the following reasons
\begin{itemize}
\item \textbf{$(A4)$} deals only with the covariance matrix and it is not clear how that alone is sufficient to say that $\bar{A}_n-(1-\beta)\mu$ is positive definite (see below Equation (26) page 16 of \cite{flstda}; the problem mentioned earlier).
\item For the sake of the argument, let us suppose that $(1-\beta)\mu$ is the smallest eigenvalue of $\bar{A}_n$. However, the problem here is that $\bar{A}_n$ is not symmetric in general and it is typical in reinforcement learning applications to say a non-symmetric matrix $M$ is positive definite if $x^\top M x>0$ for all real valued vectors $x$. While it follows from this notion of positive definiteness of $M$ that all the eigenvalues of $M$ have positive real parts, there could still be eigenvalues with complex conjugate parts. Hence to assume that the smallest eigenvalue of $\bar{A}_n$ to be $(1-\beta)\mu$ might be problematic (because there could be other complex eigenvalues with smaller magnitude). While it might be true in practice that the smallest eigenvalue of $\bar{A}_n$, is greater than $(1-\beta)\mu$ to our knowledge there is no proof for this fact in literature.
\end{itemize}
\fi
\section{Blow Up of the Bound}
We would like to note that the rate expression in Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}) contains a constant $C$. The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$.
However, it also looks like that even the constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ will blow up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved (even if we disregard all the other problems).

>>>>

\documentclass{article}
\usepackage[a4paper]{geometry}
\input{pack}
\newcommand{\Prob}[1]{\mathbb{P}\left(#1\right)}
% if you need to pass options to natbib, use, e.g.:
 \PassOptionsToPackage{compress}{natbib}
 \usepackage{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

%\usepackage{nips_2017}

@@

\title{Finite Time Bounds for Temporal Difference Learning with Function Approximation: Bugs in state-of-the-art results}

@@

\author{
Chandrashekar Lakshmi Narayanan
\and
Csaba Szepesv\'ari 
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

\maketitle
\begin{abstract}
This short report lists some problems with the recent papers by \cite{flstd,lstdicml}, whose longer version containing the proofs are available on arxiv \citep{flstda,lstdicmla}.
\end{abstract}
\section{Introduction}
This short report lists some problems with the proofs of the claims in two recent papers  by \cite{flstd,lstdicml}.
As the problems seem serious, our conclusion is that to obtain results similar to those claimed in these two works, the assumptions of the stated claims need to be considerably strengthened and the form of the results will also need to be adjusted in significant ways.

The report is not self contained and its scope is limited to an audience who are interested in RL and TD(0) and specifically the above-mentioned papers.
Furthermore, since \cite{flstd,lstdicml} omitted the proofs, we will instead discuss their longer version in what follows, which have identical statements but include the proofs. The longer version of the paper by \cite{flstd} is \citep{flstda}, while the longer version
of \cite{lstdicml} is \citep{lstdicmla}.
We will  borrow the notation directly from these works.
%While the arguments presented below are about technical issues in the analysis, we don't provide any proofs as such and hence the presentation turns out to be a commentary of the prior results.\par
%The report is organized as follows. In \Cref{sec:flstd} we discuss the issues with the approach in \cite{flstda} and in \Cref{sec:lstdicml} the roadblocks as seen in \cite{lstdicmla} are presented. Both sections have other subsections which deal with the various separate difficulties.

\section{Expected Error Bound}

\subsection{Bugs in \cite{flstda}}
One of the main results of \citet{flstda} is Theorem 1, which states a bound on the expected error.
The proof of this theorem can be found in section A.2, starting on page 16.
The proof up to Eq. (26) is correct (in the definition of $M_{n+1}$, $F_n$ should be $F$).
However, after this display, we are told that by (A4), which ensures that $\frac1T \sum_{i=1}^T \phi(s_i) \phi_t(s_i)^\top \succ \mu I$ for some $\mu>0$, $\bar A_n = \frac1n \sum_{t=1}^n \phi_t (\phi_t - \beta \phi_t')^\top$ is such that $\bar A_n - (1-\beta) \mu I $ is positive definite. 
Here, $n>0$ is an arbitrary index and for $t\ge 1$ we use the abbreviation $\phi_t = \phi(s_{i_t})$ and $\phi_t' = \phi(s_{i_t}')$, where $i_t\in \{1,\dots,T\}$ is a random index.
\begin{quote}
In general, (A4) does \emph{not} imply that  $\bar A_n - (1-\beta) \mu I $ is positive definite.
\end{quote}
Take for example $n=1$. We would need that $\phi_1 (\phi_1 - \beta \phi_1')^\top - (1-\beta) \mu I$ is positive definite.
It is easy to construct examples where this is not true: Nothing prevents, for example, $\phi_1 = \beta \phi_1'$, in which case $\bar A_1 - (1-\beta)\mu I = -(1-\beta)\mu I$ is \emph{negative} definite.
(Note that the matrices involved are \emph{not} symmetric. Unfortunately, none of the two papers defines what is meant by positive definite in this case. We assume that the definition used is that a square, real-valued matrix $A$ is called positive definite if $x^\top A x\ge 0$ for any $x$ real-valued vector of appropriate dimension.)
In fact, we don't see why the claimed relationship would hold even when 
$\bar A_n$ is replaced by $\hat A_T \doteq  \frac1T \sum_{i=1}^T \phi( s_i) (\phi(s_i) - \beta \phi(s_i'))^\top$,
and we in fact suspect that this claim is false in full generality. But at minimum, a proof would be required and the whole subsequent argument will need to be changed.
\subsection{Bugs in  \cite{lstdicmla}}
In page 10 of \cite{lstdicmla} the expression for $\mathbf{E}[a_{j+1}^\top a_{j+1}]$ contains terms that involve the product of $P$ and $P^\top$. This cannot be correct, as here we can take the expectation first over the next state, which will bring in a \emph{single} instance of $P$.
To remove clutter, drop the $j$ subindex, and set $A = \phi(\phi- \beta \phi')^\top$, where $\phi  = \phi(s_{I})$ and $\phi' = \phi(s_{I}')$.
The incriminated expression from Eq. (15) of the paper is
\begin{align}
\EE{ A - \tfrac{\gamma}{2} A^\top A  } = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(\Delta - \beta P^\top (2I - \beta \Delta)\Psi P) ) \Phi\,.
\end{align}
Here, $\Phi$ is the $S\times d$ matrix whose $i$th row is $\phi^\top(s_i)$,
$\Psi$ is the $T \times T$ diagonal matrix whose $i$th diagonal entry is $\Prob{I = i}$, while $\Delta$ is another diagonal matrix whose $i$th entry is $\norm{\phi(s_i)}_2^2$.
A direct calculation gives that
\begin{align}
A^\top A 
& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top  \nonumber \\
& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \nonumber \\
& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\, \nonumber \\
& = \norm{\phi}^2  \phi \phi^\top-\beta \norm{\phi}^2 \phi' \phi^\top -\beta \norm{\phi}^2  \phi (\phi')^\top + \beta^2 
\norm{\phi}^2 \phi' (\phi')^\top\,.
\end{align}
%where we abbreviate the 2-norm with $\norm{\cdot}$.
Hence, 
\begin{align*}
\EE{ A^\top A \,|\, I } = 
\norm{\phi}^2  \phi \phi^\top-\beta \norm{\phi}^2 \EE{ \phi' \,|\, I} \phi^\top -\beta \norm{\phi}^2  \phi (\EE{ \phi' \,|\, I})^\top + \beta^2 
\norm{\phi}^2 \EE{ \phi' (\phi')^\top\,|\, I } \,.
\end{align*}
Now, $\EE{ \phi' (\phi')^\top\,|\, I=i } = \phi(
Notice that $P \Phi$ is an $S\times d$ matrix whose $i$th row is $\EE{ \phi' \,|\, I=i}^\top$.
Hence, $ \EE{ \phi' \,|\, I} = (P \Phi)_{I:\cdot}$.
Now, $\EE{\norm{\phi}^2 \phi \phi^\top} = \Phi \Phi^\top \Psi \Phi \Phi^\top$.
Hence,
\begin{align*}
\EE{ A^\top A } = 
\EE{\norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}}\,.
\end{align*}

\section{Problems with the Proof: High Probability Bound}
\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
The proof of the high probability bound starts on page 14, in section A.1.
The first problem happens in the display on the bottom of this page in the proof of Lemma 6.
Here, we are told that for $A = \phi \phi^\top - \beta \phi (\phi')^\top$ (we are dropping indices to remove clutter),
\begin{align*}
A^\top A = \norm{\phi}^2 \phi \phi^\top - \beta ( 2 - \norm{\phi}^2\beta ) \phi' (\phi')^\top\,,
\end{align*}
where $\norm{x}$ denotes the $2$-norm of $x$.
However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation gives:
\begin{align*}
A^\top A 
& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top \\
& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \\
& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\,,
\end{align*}
which does not match the previous display.
The terms that do not match are the linear-in-$\beta$ terms.
In the first display we have $ -2 \beta \phi' (\phi')^\top$, while in the bottom we have $-\beta (\phi' \phi^\top + \phi (\phi')^\top)$.

The first equality of their display states (in equivalent form) that 
\begin{align*}
A^\top A = \norm{\phi}^2 ( \phi \phi^\top - 2\beta \phi (\phi')^\top +\beta^2 \phi' (\phi')^\top )\,.
\end{align*}
We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.
\begin{comment}
We also have a hard time following the rest of the proof of this lemma.
In fact, we have a simple proof for a result similar to stated in this lemma.
This works by observing that one needs to 
bound the Lipschitz factor of $\theta \mapsto \bar T_n \dots \bar T_{t+1} ( (I-\gamma_t A_t)\theta + b_t )$, where
$A_t = \phi_t (\phi_t -\beta \phi_t')^\top$, $b_t = r_{i_t} \phi_t$ and
 $\bar T_t(\theta) = (I-\gamma_t \bar A_T) \theta + \bar b_T$ (we assume that the stepsize sequence is deterministic).
 Clearly, this Lipschitz factor can be bounded by $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 \norm{I-\gamma_t A_t}_2$
 (the composition of Lipschitz maps is Lipschitz with a factor that is the product of the individual maps' Lipschitz factors). 
Invoking (A3) then indeed gives some version of Lemma 7.
By making an appropriate assumption (similar to those used in our paper), 
$\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.
\end{comment}
\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
In page 14 the expression below "A is a possibly random matrix..." is not entirely justified (personal communication with one of the authors).

\section{Bounded Linear Space Assumption}
\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
\citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}). This assumption is critical in establishing Lemma~7 (see pages 15 and 16, \cite{flstda}).
We suspect the following shortcomings with assumption \textbf{$(A3)$}
\begin{itemize}
\item The assumption is stated in a somewhat sloppy fashion. We take the authors meant to say that $\sup_n \norm{\Phi \theta_n}_\infty<+\infty$ holds almost surely. This seems like a strong assumption! Would not one of the main question be whether this assumption even holds?
% We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
\item One possibility would be to modify the algorithm by adding a projection step to guarantee boundedness.
It is still unclear  whether this alone would ensure convergence of the error to zero. In any case, the expected error bound analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).

%For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly). Then would the recursion unfold the same way as it does in Equation (26) of \cite{flstda} (or Equation (13) of \cite{flstd})? How are we to ensure decay/contraction (as shown in Equation (27) of \cite{flstda}) at each step (since projection might be non-contracting)?
%\item The recursion eventually (see Equation (26) of \cite{flstda}) has only \emph{additive noise} and not \emph{multiplicative noise} as analyzed in our submission. The additive noise is captured in the form of the $H_\beta^2$ term in Equation (28) of \cite{flstd}. This reduction to additive noise is possible due to \textbf{$(A3)$}.
\end{itemize}
\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
\citet{gugan} mention that (citing a personal communication with \citet{lstdicmla}) \cite{lstdicmla} assume implicitly the projection step in all the high probability bounds in \cite{lstdicmla}. While this implicit projection in itself does not affect the high probability bound proofs directly, it is important to note that the set to which the iterates needs to be projected has to contain the TD(0) solution, a condition that cannot be guaranteed.

\section{Relation between Covariance Matrix and $\bar{A}_T$ matrix}
\citet{flstda} assume positive definiteness of \textbf{$(A4)$} covariance matrix $\frac{1}{T}\Phi^\top_T \Phi_T$. However, unlike regression problems, in reinforcement learning problems what appears in the recursion (see Equation (6)) is not the covariance matrix, but a different matrix $\bar{A}_T=\frac{1}{T}\sum_{i=1}^T \phi(s_i)(\phi(s_i)-\beta\phi(s_i'))^\top$ defined in pages 2, 4 (below Equation (5)), 8 and 16 of \cite{flstda}. Usually, without a sampling assumption known as the `on-policy' case (see \cite{gtd} for a discussion on `on-policy' vs `off-policy') the eigenvalues of $\bar{A}_T$ cannot be guaranteed to have all positive real parts. While \citet{flstda} mention the `on-policy' sampling in the introduction, there is no explicit sampling assumption in the list of assumption. In fact, we doubt that the proposed algorithm will converge without extra assumption (as discussed above).
\if0
This is worrisome for the following reasons
\begin{itemize}
\item \textbf{$(A4)$} deals only with the covariance matrix and it is not clear how that alone is sufficient to say that $\bar{A}_n-(1-\beta)\mu$ is positive definite (see below Equation (26) page 16 of \cite{flstda}; the problem mentioned earlier).
\item For the sake of the argument, let us suppose that $(1-\beta)\mu$ is the smallest eigenvalue of $\bar{A}_n$. However, the problem here is that $\bar{A}_n$ is not symmetric in general and it is typical in reinforcement learning applications to say a non-symmetric matrix $M$ is positive definite if $x^\top M x>0$ for all real valued vectors $x$. While it follows from this notion of positive definiteness of $M$ that all the eigenvalues of $M$ have positive real parts, there could still be eigenvalues with complex conjugate parts. Hence to assume that the smallest eigenvalue of $\bar{A}_n$ to be $(1-\beta)\mu$ might be problematic (because there could be other complex eigenvalues with smaller magnitude). While it might be true in practice that the smallest eigenvalue of $\bar{A}_n$, is greater than $(1-\beta)\mu$ to our knowledge there is no proof for this fact in literature.
\end{itemize}
\fi
\section{Blow Up of the Bound}
We would like to note that the rate expression in Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}) contains a constant $C$. The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$.
However, it also looks like that even the constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ will blow up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved (even if we disregard all the other problems).
